---
title: "homework-week-4"
author: "Nitin Verma"
date: "September 29, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Homework Problem 1:
Every Saturday, at the same time, a primatologist goes and sits in the forest in the morning and listens for titi monkey calls, counting the number of calls they hear in a 2 hour window from 5am to 7am. Based on previous knowledge, she believes that the mean number calls she will hear in that time is exactly 15. Let X represent the appropriate Poisson random variable of the number of calls heard in each monitoring session.
- What is the probability that she will hear more than 8 calls during any given session?
```{r}
l <- 15
x <- 8

p <- 1 - ppois(q = x, lambda = l)
p
```
- What is the probability that she will hear no calls in a session?
```{r}
p <- dpois(x = 0, lambda = l)
p
```
- What is the probability that she will hear exactly 3 calls in a session?
```{r}
p <- dpois(x = 3, lambda = l)
p
```
- Plot the relevant Poisson mass function over the values in range 0 ≤ x ≤ 30.
```{r}
x <- 0:30
l <- 15
probset <- dpois(x = x, lambda = l)
barplot(probset, names.arg = x, space = 0, xlab = "x", ylab = "Pr(X = x)", main = "Probability Mass Function")
```
<br>
- Simulate 104 results from this distribution (i.e., 2 years of Saturday monitoring sessions).
```{r}
x <- rpois(104, lambda = l)
x
fx <- (l^x * exp(-l))/factorial(x)
fx
```
- Plot the simulated results using hist() and use xlim() to set the horizontal limits to be from 0 to 30. How does your histogram compare to the shape of the probability mass function you plotted above?
```{r}
hist(x = x, xlim = c(0,30), freq = FALSE, col = "light blue", xlab = "No. of Titi Monkey Calls per session", ylab = "Probability", main = "Histogram of Titi Monkey Calls")
```

<em><b>Comparison between the histogram and the PMF:</b></em><br>
In terms of upper and lower limits, both plots appear to be restricted in the interval (0, 30). Also, the histogram seems to have a mode around 15 calls per session. The histogram appears broadly to be the same shape as the PMF.
<br>
<br>
